---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# exceldata

<!-- badges: start -->
[![Lifecycle: Stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/exceldata)](https://CRAN.R-project.org/package=exceldata)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/grand-total/exceldata)](https://cran.r-project.org/package=exceldata)
[![R-CMD-check](https://github.com/biostatsPMH/exceldata/workflows/R-CMD-check/badge.svg)](https://github.com/biostatsPMH/exceldata/actions)
<!-- badges: end -->

The goal of exceldata is to facilitate the use of Excel as a data entry tool for reproducible research. This package provides tools to automate data cleaning and recoding of data by requiring a data dictionary to accompany the Excel data. A macro-enabled template file has been created to facilitate clean data entry including validation rules for live data checking.

To download the Excel Template
[Click Here](https://github.com/biostatsPMH/exceldata/blob/main/images/DataDictionary0.9.1.xlsm) and then click on `Download`

To view data in the required format without macros
[Click Here](https://github.com/biostatsPMH/exceldata/blob/main/inst/extdata/exampleData.xlsx) and then click on `Download`

## Installation

You can install the released version of exceldata from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("exceldata")
```

And the development version from [GitHub](https://github.com/) with:

``` r
devtools::install_github("biostatsPMH/exceldata")
```

---
## Documentation


[Online Documentation](https://biostatsPMH.github.io/exceldata/)

PDF Documentation: [Click Here](https://github.com/biostatsPMH/exceldata/blob/main/docs/ExcelDictionaryUserManual.pdf) and then click on `Download`
<!-- Note - this is created in a separate directory - Teaching/excelData Instructions -->

[Using the Excel Template](https://biostatsPMH.github.io/exceldata/data-dictionary-1.html)


## Example

Example of importing the data and producing univariate plots to screen for outliers.

```{r example}
library(exceldata)

exampleDataFile <- system.file("extdata", "exampleData.xlsx", package = "exceldata")
import <- importExcelData(exampleDataFile,dictionarySheet = 'DataDictionary',dataSheet = 'DataEntry')

# The imported data dictionary 
dictionary <- import$dictionary
head(dictionary)

# The imported data, with calculated variables
data <- import$data
head(data)

# Simple univariate plots with outliers 
plots <- plotVariables(data,dictionary,IDvar = 'ID',showOutliers = T)
# Not Run: Show plots in viewer
# plots

```

